---
title: 'Tidy Tuesday for 9/1/2020'
author: " "
date: " "
output: html_document
---

For this Tidy Tuesday challenge we used data files from *Our World in Data* describing crop yields across the globe from 1960-present.  Data can be found at https://ourworldindata.org/crop-yields

Peas seemed appealing at the time and I never see them growing here in Maryland-- where do they come from?

I plotted a line graph with tons of peas produced per hectare on the y-axis, year (range from 1990 to 2020) on the x-axis, and continent as line color.

I learned how to use the directlabels package to put continent names next to lines rather than using a legend connecting color to continent.

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load libaries
library (tidyverse)
library (directlabels)
library (patchwork)

#set wd
setwd ("/Users/jprout/OneDrive - University of Maryland School of Medicine/R/TidyTues/crops")

#Load data
crop <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-01/key_crop_yields.csv')

#Change colname 
colnames(crop)[10] <- "peas"

#subset to peas
peas <- select (crop, Year, peas, Entity, Code)

#make NA equal Region/Continent for code
peas$Code <- fct_explicit_na(peas$Code, na_level = "Region/Continent")

#change Northern America to North America
peas$Entity <- recode (peas$Entity, "Northern America" = "North America")


#filter just to include continents
p_cont <- filter (peas, Entity == "Africa" | Entity == "Asia"|
                    Entity == "Australia"| Entity == "Europe"|
                    Entity == "South America"| Entity == "North America")


#plot amount of peas by year and entity for continents
continent_plot <- ggplot (data = p_cont, aes (x=Year, y = peas, color = Entity)) + 
  geom_smooth (se=FALSE)+
  theme (axis.text.x = element_text (color = "black"), 
         axis.text.y = element_text (color = "black"),  
         panel.background = element_blank (),
         axis.line = element_line (color = "black"),
         legend.title = element_blank(), 
         legend.position="none",
         plot.margin = margin (0.1, 2.6, 0.1, 0.1, "cm"))+
  labs(title = "North America Takes Lead in Pea Production", 
       x = " ", 
       y = "Tons per Hectare")  

#adding labels at end
continent_plot_labels <- direct.label (continent_plot, method = list(col = "black", "last.bumpup")) +
  scale_y_continuous (limits = c(0,3))+
  scale_x_continuous (limits = c (1990, 2020))+
  coord_cartesian (clip = "off")

print (continent_plot_labels)

```



This graph was more interesting than I thought it would be.  Why did North American pea production increase while European pea production dipped lower.  To look at this question I plotted pea production over time for Europe, North America, and Canada.


```{r Peas in Europe and North America, warning= FALSE, include= TRUE, message= FALSE}
#getting North American and European divisions (dont see Western Europe)
EuropeandNA <- filter (peas, Entity == "Europe" |Entity == "United States"|
                    Entity == "Canada")

#limiting time frame from 1990-present
EuropeandNA <- filter (EuropeandNA, Year >= 1990)

#line chart of pea production over time in Europe and NA
europeNA <- ggplot (data = EuropeandNA, aes (x=Year, y = peas, color = Entity)) + 
  geom_smooth (se=FALSE) +
  theme (axis.text.x = element_text (color = "black"), 
         axis.text.y = element_text (color = "black"),  
         panel.background = element_blank (),
         axis.line = element_line (color = "black"),
         legend.title = element_blank(), 
         legend.position="none",
         plot.margin = margin (0.1, 2.6, 0.1, 0.1, "cm")) +
  labs(title = "Canada Drives Increase in North American Peas", 
       x = " ", 
       y = "Tons per Hectare") 

europeNAwithlabels <- direct.label (europeNA, method = list(col = "black", "last.bumpup")) +
  scale_y_continuous (limits = c(0,3))+
  scale_x_continuous (limits = c (1990, 2020))+
  coord_cartesian (clip = "off")

print (europeNAwithlabels)
```

Looks like more peas are coming from Canada!  They must like cold weather.

My other goal was to use the patchwork package to combine graphs.

```{r patchwork, include = TRUE, message=FALSE, warning = FALSE}
patchwork <- continent_plot_labels + europeNAwithlabels+ plot_layout (ncol = 1)

print(patchwork)
```

